---
title: "Final Project Peer Review Draft"
author: "Molly Robertson"
date: "4/12/2020"
output: html_document
---

#Final Project Peer Review Draft 

Topic Overview: As an immunology and infectious disease major, I am interested in looking at various countries response to coronavirus from December 2019 to current. I would look at the total number of cases as they grew, as well as the response measurements and then the decline. Specifically, I would want to see which measures had the greatest impact on slowing down the case numbers, and identify which countries did a good job.  

```{r} 
#Load In Necessary Packages
library(DataComputing)

```

###Loading In Data 
```{r}
#Rename Data Tables
GlobalRecovered <- read.csv("https://data.humdata.org/hxlproxy/data/download/time_series_covid19_recovered_global_iso3_regions.csv?dest=data_edit&filter01=merge&merge-url01=https%3A%2F%2Fdocs.google.com%2Fspreadsheets%2Fd%2Fe%2F2PACX-1vTglKQRXpkKSErDiWG6ycqEth32MY0reMuVGhaslImLjfuLU0EUgyyu2e-3vKDArjqGX7dXEBV8FJ4f%2Fpub%3Fgid%3D1326629740%26single%3Dtrue%26output%3Dcsv&merge-keys01=%23country%2Bname&merge-tags01=%23country%2Bcode%2C%23region%2Bmain%2Bcode%2C%23region%2Bmain%2Bname%2C%23region%2Bsub%2Bcode%2C%23region%2Bsub%2Bname%2C%23region%2Bintermediate%2Bcode%2C%23region%2Bintermediate%2Bname&filter02=merge&merge-url02=https%3A%2F%2Fdocs.google.com%2Fspreadsheets%2Fd%2Fe%2F2PACX-1vTglKQRXpkKSErDiWG6ycqEth32MY0reMuVGhaslImLjfuLU0EUgyyu2e-3vKDArjqGX7dXEBV8FJ4f%2Fpub%3Fgid%3D398158223%26single%3Dtrue%26output%3Dcsv&merge-keys02=%23adm1%2Bname&merge-tags02=%23country%2Bcode%2C%23region%2Bmain%2Bcode%2C%23region%2Bmain%2Bname%2C%23region%2Bsub%2Bcode%2C%23region%2Bsub%2Bname%2C%23region%2Bintermediate%2Bcode%2C%23region%2Bintermediate%2Bname&merge-replace02=on&merge-overwrite02=on&tagger-match-all=on&tagger-01-header=province%2Fstate&tagger-01-tag=%23adm1%2Bname&tagger-02-header=country%2Fregion&tagger-02-tag=%23country%2Bname&tagger-03-header=lat&tagger-03-tag=%23geo%2Blat&tagger-04-header=long&tagger-04-tag=%23geo%2Blon&header-row=1&url=https%3A%2F%2Fraw.githubusercontent.com%2FCSSEGISandData%2FCOVID-19%2Fmaster%2Fcsse_covid_19_data%2Fcsse_covid_19_time_series%2Ftime_series_covid19_recovered_global.csv")
GlobalDeaths<- read.csv("https://data.humdata.org/hxlproxy/data/download/time_series_covid19_deaths_global_iso3_regions.csv?dest=data_edit&filter01=merge&merge-url01=https%3A%2F%2Fdocs.google.com%2Fspreadsheets%2Fd%2Fe%2F2PACX-1vTglKQRXpkKSErDiWG6ycqEth32MY0reMuVGhaslImLjfuLU0EUgyyu2e-3vKDArjqGX7dXEBV8FJ4f%2Fpub%3Fgid%3D1326629740%26single%3Dtrue%26output%3Dcsv&merge-keys01=%23country%2Bname&merge-tags01=%23country%2Bcode%2C%23region%2Bmain%2Bcode%2C%23region%2Bmain%2Bname%2C%23region%2Bsub%2Bcode%2C%23region%2Bsub%2Bname%2C%23region%2Bintermediate%2Bcode%2C%23region%2Bintermediate%2Bname&filter02=merge&merge-url02=https%3A%2F%2Fdocs.google.com%2Fspreadsheets%2Fd%2Fe%2F2PACX-1vTglKQRXpkKSErDiWG6ycqEth32MY0reMuVGhaslImLjfuLU0EUgyyu2e-3vKDArjqGX7dXEBV8FJ4f%2Fpub%3Fgid%3D398158223%26single%3Dtrue%26output%3Dcsv&merge-keys02=%23adm1%2Bname&merge-tags02=%23country%2Bcode%2C%23region%2Bmain%2Bcode%2C%23region%2Bmain%2Bname%2C%23region%2Bsub%2Bcode%2C%23region%2Bsub%2Bname%2C%23region%2Bintermediate%2Bcode%2C%23region%2Bintermediate%2Bname&merge-replace02=on&merge-overwrite02=on&tagger-match-all=on&tagger-01-header=province%2Fstate&tagger-01-tag=%23adm1%2Bname&tagger-02-header=country%2Fregion&tagger-02-tag=%23country%2Bname&tagger-03-header=lat&tagger-03-tag=%23geo%2Blat&tagger-04-header=long&tagger-04-tag=%23geo%2Blon&header-row=1&url=https%3A%2F%2Fraw.githubusercontent.com%2FCSSEGISandData%2FCOVID-19%2Fmaster%2Fcsse_covid_19_data%2Fcsse_covid_19_time_series%2Ftime_series_covid19_deaths_global.csv")
GlobalConfirmed <- read.csv("https://data.humdata.org/hxlproxy/data/download/time_series_covid19_confirmed_global_iso3_regions.csv?dest=data_edit&filter01=merge&merge-url01=https%3A%2F%2Fdocs.google.com%2Fspreadsheets%2Fd%2Fe%2F2PACX-1vTglKQRXpkKSErDiWG6ycqEth32MY0reMuVGhaslImLjfuLU0EUgyyu2e-3vKDArjqGX7dXEBV8FJ4f%2Fpub%3Fgid%3D1326629740%26single%3Dtrue%26output%3Dcsv&merge-keys01=%23country%2Bname&merge-tags01=%23country%2Bcode%2C%23region%2Bmain%2Bcode%2C%23region%2Bmain%2Bname%2C%23region%2Bsub%2Bcode%2C%23region%2Bsub%2Bname%2C%23region%2Bintermediate%2Bcode%2C%23region%2Bintermediate%2Bname&filter02=merge&merge-url02=https%3A%2F%2Fdocs.google.com%2Fspreadsheets%2Fd%2Fe%2F2PACX-1vTglKQRXpkKSErDiWG6ycqEth32MY0reMuVGhaslImLjfuLU0EUgyyu2e-3vKDArjqGX7dXEBV8FJ4f%2Fpub%3Fgid%3D398158223%26single%3Dtrue%26output%3Dcsv&merge-keys02=%23adm1%2Bname&merge-tags02=%23country%2Bcode%2C%23region%2Bmain%2Bcode%2C%23region%2Bmain%2Bname%2C%23region%2Bsub%2Bcode%2C%23region%2Bsub%2Bname%2C%23region%2Bintermediate%2Bcode%2C%23region%2Bintermediate%2Bname&merge-replace02=on&merge-overwrite02=on&tagger-match-all=on&tagger-01-header=province%2Fstate&tagger-01-tag=%23adm1%2Bname&tagger-02-header=country%2Fregion&tagger-02-tag=%23country%2Bname&tagger-03-header=lat&tagger-03-tag=%23geo%2Blat&tagger-04-header=long&tagger-04-tag=%23geo%2Blon&header-row=1&url=https%3A%2F%2Fraw.githubusercontent.com%2FCSSEGISandData%2FCOVID-19%2Fmaster%2Fcsse_covid_19_data%2Fcsse_covid_19_time_series%2Ftime_series_covid19_confirmed_global.csv")
```


###Looking at Grand Totals by Worldwide Region
At the most basic level, I wanted to look at the total confirmed cases, deaths, and recovered based on regions. The regions in the dataset are Africa, Americas, Asia, Europe and Oceania. 
First, is the table and graph for confirmed cases, followed by deaths and then recovered. 
```{r}
GlobalConfirmed1<-
GlobalConfirmed%>%
  group_by(Region.Name)%>%
  summarise(totalconfirmed = sum((X4.11.20)))
GlobalConfirmed1<-
 GlobalConfirmed1[-2,]
GlobalConfirmed1<-
  GlobalConfirmed1[-1,]
head(GlobalConfirmed1)
```

```{r}
GlobalConfirmed1%>%
  ggplot(aes(x=Region.Name, y=totalconfirmed)) +
  geom_bar(stat = "identity",aes(color=Region.Name, fill=Region.Name)) +
  ggtitle("Total Confirmed Cases by Region") +
  ylim(0, 875000)
```

Looking at the graph above, it appears as though Europe has the most cases with the Americas behind. These graphs need to be further divided to get the entire picture, as the majority of cases in the Americas are in the United States. Certain regions are dominated by certains countries. 


```{r}
GlobalDeaths1<-
GlobalDeaths%>%
  group_by(Region.Name)%>%
  summarise(Total_Deaths = sum((X4.11.20)))
GlobalDeaths1<-
 GlobalDeaths1[-2,] #get rid of useless row
GlobalDeaths1<-
  GlobalDeaths1[-1,] #get rid of useless row 
head(GlobalDeaths1)
```

```{r}
GlobalDeaths1%>%
  ggplot(aes(x=Region.Name, y=Total_Deaths)) +
  geom_bar(stat = "identity",aes(color=Region.Name, fill=Region.Name)) +
  ggtitle("Total Deaths by Region") +
  ylim(0, 75000)
```

Unfortunately, Europe has the highest total number of deaths as well. It must be noted that many European countries are a few weeks ahead of the outbreak in the Americas, specifically the United States so this can change. These graphs also do not take into account total proportion of a population that have been infected, these are just pure counts. 

```{r}
GlobalRecovered1<-
GlobalRecovered%>%
  group_by(Region.Name)%>%
  summarise(Total_Recovered = sum((X4.11.20)))
GlobalRecovered1<-
 GlobalRecovered1[-2,]
GlobalRecovered1<-
  GlobalRecovered1[-1,]
head(GlobalRecovered1)
```

```{r}
GlobalRecovered1%>%
  ggplot(aes(x=Region.Name, y=Total_Recovered)) +
  geom_bar(stat = "identity",aes(color=Region.Name, fill=Region.Name)) +
  ggtitle("Total Recovered by Region")
```

Upliftingly, Europe does also have the highest amount of total recovered. Sadly, the Americas is still low on the recovered as we are nearing the peak of the outbreak in the United States. If the data we currently have is correct, the Asia region was able to contain their outbreak, mostly China, and keep deaths relatively low with recoveries high. 


These graphs are fairly basic, and do not give us great insight into the outbreak. To further investigate this pandemic, it is necessary to look at a few countries from each of the regions, and their specific plan to combat the spread of the coronavirus. 





















